---
title: "Microstats vignette"
output:
  html_document:
    df_print: paged
vignette: "%\\VignetteIndexEntry{Microstats Vignette} %\\VignetteEngine{knitr::rmarkdown}
  %\\VignetteEncoding{UTF-8}\n"
---

```{r setup, message=FALSE, echo=FALSE}
library(knitr)
```

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

```{r, echo = FALSE}
library(MicroStats)
```

### Introduction to <i>Microstats</i>

The <i>Microstats</i> package allows users to analyze microbial growth data into quantifiable values and images.


### <i>Microstats</i> Function Overview
*`growthRates`: used to calculate, plot, and test microbial data.


### Required Software
This program requires installation R (and recommends RStudio) but prior R knowledge is not necessary. To use this package, it must first be installed and loaded into your R session. All functions can be further explored by using `?functionName`, where additional information for the function and its parameters can be found.


### Preparing Data
`growthRates` requires two dataframes. One containing the optical density (OD) values or raw with the first column as time, this dataframe may be referred to as `d` throughout the package. The second dataframe should contain metadata about the wells or map data, such as strain type and environment. This dataframe may be referred to as `wells`. Of course, the analysis is only as accurate as the data collected. 

<b>Important</b> The `d` dataframe's first column must be time!

Additionally, the columns of `d` should match with the rows of `wells` such that the first row in wells and the first column of data (column 2) of d should match. Within the functions, data will be cleaned and columns renamed, so manual cleaning is not required. In function cleaning consists of considering only the complete cases with respect to the time variable.

### Examples of Dataframes
(Put images here)


### Running growthRates
The growthRates function is used to analyze two sets of information (OD values/time and well information) and calculate the growth information of the given data (log growth curve, maximum optical density). 

This function can be used either in a file (.R, .Rmd, etc.) or in the console. If the function is being used in the console or an R file, any plots will appear in the `plots` pane if you are using RStudio, and all output will be printed to the console. Once you run the function, it will save a dataframe called `data` to R. This dataframe will consist of the original well information, the maximum slope of LOESS-smoothed log growth curve (lgr), and the maximum optical density (maxOD) for the well.

growthRates takes in two dataframes (`d` and `wells`) as well as logical and string values indicating which output you are looking for and any additional information. The first of which, `plotRaw` indicates whether you would like to have your raw data plotted.By default, plotRaw is set to true, but can be changed to false by indicating `plotRaw = FALSE`. If you would like it to remain true, you can omit setting this parameter when calling the function. The second parameter is `strainName` which takes in a string value of the strain's name.

To run the function, type in the function name and the specified arguments like so:
```r
growthRates(myODData, myWellData, plotRaw = FALSE, strainName = "Strain's name")
```
Once you have run the function, you should find the `data` dataframe saved (to your global environment if you are using RStudio).

If you have attempted to plot the outputs, but it results in an error or doesn't plot, enlarge your plots pane as the cause is likely the image is too large for the plots pane. Adjusting the size of the plots pane should resolve this issue.


